---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# coasts

<!-- badges: start -->
[![R-CMD-check](https://github.com/WorldFishCenter/peskas.coasts/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/WorldFishCenter/peskas.coasts/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/WorldFishCenter/peskas.coasts/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/WorldFishCenter/peskas.coasts/actions/workflows/pkgdown.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
<!-- badges: end -->

The `coasts` package provides a comprehensive data pipeline for processing and analyzing coastal fisheries data from the Western Indian Ocean (WIO) region. It integrates GPS tracking data from Pelagic Data Systems (PDS), regional fisheries metadata, and geospatial information to support fisheries research and management.

## Overview

This package is designed to handle the complete workflow for coastal fisheries data processing:

1. **Data Ingestion**: Automated retrieval of GPS boat tracking data from Pelagic Data Systems
2. **Data Preprocessing**: Spatial gridding and summarization of fishing activity patterns  
3. **Data Export**: Integration with MongoDB for data storage and geospatial analysis
4. **Metadata Management**: Handling of device information and regional boundaries

The package supports data from Kenya and Zanzibar fisheries, with built-in currency conversion and regional harmonization capabilities.

## Key Features

- **GPS Track Processing**: Ingest and preprocess boat GPS tracks from PDS API
- **Spatial Analysis**: Grid-based summarization of fishing activity at multiple scales (100m-1km)
- **Cloud Storage Integration**: Seamless upload/download from Google Cloud Storage with multi-bucket support for regional data
- **MongoDB Integration**: Geospatial data storage with 2dsphere indexing
- **Parallel Processing**: Efficient handling of large datasets using parallel computation
- **Automated Pipeline**: GitHub Actions workflow for continuous data processing

## Installation

You can install the development version of coasts from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("WorldFishCenter/peskas.coasts")
```

## Configuration

### Local Development Setup

For local development, the package uses environment variables managed through a `.env` file:

1. Copy the `.env.example` file to `.env`:
   ```bash
   cp .env.example .env
   ```

2. Fill in your credentials in the `.env` file. Required environment variables include:
   - `PDS_TOKEN`, `PDS_SECRET`, `PDS_USERNAME`, `PDS_PASSWORD`, `PDS_CUSTOMER_ID`: Pelagic Data Systems API credentials
   - `GCP_SA_KEY`: Google Cloud service account key (JSON format)
   - `MONGODB_CONNECTION_STRING_COASTS`, `MONGODB_CONNECTION_STRING_TRACKS`: MongoDB connection strings
   - `GOOGLE_SHEET_ID`: Google Sheets ID for metadata
   - `AIRTABLE_TOKEN`, `AIRTABLE_BASE_ID_FRAME`, `AIRTABLE_BASE_ID_TRACKS_APP`: Airtable credentials

The package automatically loads these environment variables when running locally.

### Production Deployment

For production environments, set these environment variables directly in your deployment configuration (e.g., GitHub Secrets, Docker environment, etc.). The package will use them automatically without requiring a `.env` file.

## Main Functions

### Data Ingestion

```{r eval=FALSE}
library(coasts)

# Ingest GPS trip data from PDS
ingest_pds_trips()

# Ingest detailed GPS track data
ingest_pds_tracks()
```

### Data Preprocessing

```{r eval=FALSE}
# Preprocess tracks into spatial grids
preprocess_pds_tracks(grid_size = 500)  # 500m grid cells

# Available grid sizes: 100, 250, 500, 1000 meters
preprocess_pds_tracks(grid_size = 1000)  # 1km grid cells
```

### Data Export

```{r eval=FALSE}
# Export processed data to MongoDB
export_geos()
```

### Metadata Management

```{r eval=FALSE}
# Get device metadata from Google Sheets
devices <- get_metadata(table = "devices")

# Get all metadata tables
all_metadata <- get_metadata()
```

## Data Pipeline Workflow

The package implements an automated data pipeline that runs every 2 days via GitHub Actions:

1. **Build Container**: Creates a Docker container with R and all dependencies
2. **Ingest PDS Trips**: Downloads trip metadata from PDS API
3. **Ingest PDS Tracks**: Downloads detailed GPS tracks for each trip
4. **Preprocess Tracks**: Creates spatial grid summaries of fishing activity
5. **Export Data**: Uploads processed data to MongoDB with geospatial indexing

## Data Products

The pipeline produces several key data products:

- **Trip Data**: Basic trip information (start/end times, vessel info)
- **Track Data**: Detailed GPS points with speed, heading, and temporal information
- **Grid Summaries**: Spatial aggregations showing:
  - Time spent fishing in each grid cell
  - Average speed and vessel metrics
  - Temporal patterns of activity
- **Regional Metrics**: Time series data with currency-converted economic indicators

## Spatial Analysis Capabilities

The package supports multi-scale spatial analysis:

```{r eval=FALSE}
# Fine-scale analysis (100m grids)
preprocess_pds_tracks(grid_size = 100)

# Broad-scale patterns (1km grids)  
preprocess_pds_tracks(grid_size = 1000)
```

Grid summaries include:
- `time_spent_mins`: Total fishing time per grid cell
- `mean_speed`: Average vessel speed
- `n_points`: Number of GPS observations
- `first_seen`/`last_seen`: Temporal extent of activity

## Cloud Storage Integration

The package seamlessly integrates with Google Cloud Storage and supports multiple buckets for regional/environment-specific data:

```{r eval=FALSE}
# List available buckets
list_storage_buckets()
# Returns: c("kenya-prod", "mozambique-prod", "zanzibar-prod", "dev", "prod")

# Download from default bucket (backward compatible)
raw_data <- download_parquet_from_cloud(
  prefix = "pds-trips",
  provider = conf$storage$google$key,
  options = conf$storage$google$options
)

# Download from a specific regional bucket
kenya_data <- download_parquet_from_cloud(
  prefix = "pds-trips",
  provider = conf$storage$google$key,
  options = conf$storage$google$options,
  bucket_name = "kenya-prod"
)

# Upload to a specific regional bucket
upload_parquet_to_cloud(
  data = processed_data,
  prefix = "pds-trips-processed",
  provider = conf$storage$google$key,
  options = conf$storage$google$options,
  bucket_name = "mozambique-prod"
)

# Get bucket configuration
bucket_config <- get_bucket_config("zanzibar-prod")
```

## MongoDB Integration

Geospatial data is stored in MongoDB with appropriate indexing:

```{r eval=FALSE}
# Push data with geospatial indexing
mdb_collection_push(
  data = spatial_data,
  connection_string = "mongodb://...",
  collection_name = "fishing_areas",
  geo = TRUE  # Creates 2dsphere index
)
```

## Contributing

This package is part of the WorldFish Center's Peskas initiative for small-scale fisheries monitoring. Contributions are welcome via GitHub issues and pull requests.

## License

GPL (>= 3)
