---
title: "Peskas Tracks Usage -- Analytics Report"
date: "`r format(params$report_date, '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    self_contained: true
params:
  catch_record_usage: NULL
  usage_summary: NULL
  top_n: 50
  report_date: !expr Sys.Date()
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 14,
  fig.height = 8
)

library(dplyr)
library(ggplot2)
library(reactable)
```

```{css, echo=FALSE}
.main-container {
  max-width: 95% !important;
  margin-left: auto;
  margin-right: auto;
}

body {
  max-width: 95% !important;
}
```

## Executive Summary

This report provides an overview of app usage (recorded catches by users) across Kenya, Zanzibar, and Mozambique.

```{r summary_stats}
total_boats <- dplyr::n_distinct(params$usage_summary$imei)
total_catches <- sum(params$usage_summary$total_records)
active_countries <- dplyr::n_distinct(params$usage_summary$country)

avg_catches_per_boat <- round(mean(params$usage_summary$total_records), 1)
```

**Key Metrics:**

-   **Total Active Boats:** `r total_boats`
-   **Total Catch Records:** `r total_catches`
-   **Countries:** `r active_countries`
-   **Average Catches per Boat:** `r avg_catches_per_boat`

------------------------------------------------------------------------

## Usage Heatmap

The heatmap below shows catch recording activity over time, with boats ranked by total usage (highest at top).

```{r heatmap, fig.height=10}
# Create heatmap
coasts::create_usage_heatmap(params$catch_record_usage, top_n = params$top_n)
```

------------------------------------------------------------------------

## Usage by Country

```{r country_summary}
country_summary <- params$usage_summary |>
  dplyr::group_by(country) |>
  dplyr::summarise(
    boats = dplyr::n(),
    total_catches = sum(total_records),
    avg_per_boat = round(mean(total_records), 1),
    .groups = "drop"
  ) |>
  dplyr::arrange(dplyr::desc(total_catches))

reactable::reactable(
  country_summary,
  columns = list(
    country = reactable::colDef(name = "Country", minWidth = 150),
    boats = reactable::colDef(name = "Active Boats", align = "center"),
    total_catches = reactable::colDef(name = "Total Catches", align = "center"),
    avg_per_boat = reactable::colDef(name = "Avg per Boat", align = "center")
  ),
  striped = TRUE,
  bordered = TRUE,
  highlight = TRUE
)
```

------------------------------------------------------------------------

## Detailed Usage Table

Searchable and filterable table showing all boats and their usage statistics.

```{r usage_table}
# Create interactive table
coasts::create_usage_table(params$usage_summary)
```

------------------------------------------------------------------------

## Top Performers

```{r top_performers}
top_10 <- params$usage_summary |>
  dplyr::slice_head(n = 10)

reactable::reactable(
  top_10,
  columns = list(
    country = reactable::colDef(name = "Country", width = 100),
    boat_name = reactable::colDef(name = "Boat", minWidth = 150),
    total_records = reactable::colDef(
      name = "Total Catches",
      align = "center",
      style = list(fontWeight = 700, color = "#2e7d32")
    ),
    weeks_active = reactable::colDef(name = "Weeks Active", align = "center"),
    avg_records_per_week = reactable::colDef(
      name = "Avg/Week",
      align = "center",
      format = reactable::colFormat(digits = 1)
    )
  ),
  defaultColDef = reactable::colDef(
    headerStyle = list(background = "#f5f5f5", fontWeight = 700)
  ),
  striped = TRUE,
  bordered = TRUE
)
```

------------------------------------------------------------------------

::: {style="text-align: center; margin-top: 40px; color: #666; font-size: 0.9em;"}
Report generated on `r format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')`
:::